---
title: "Homework 2"
author: "Mayuri Albal"
date: "2022-09-30"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(readxl)
```

#Problem 1

###Cleaning
```{r}
nyc_transit=read_csv('/Users/mayurialbal/Documents/Data Science/Homework/HW2/HW2/NYC_Transit_Subway_Entrance_And_Exit_Data.csv',
      col_types=cols(Route8 ='c', Route9 ='c', Route10 ='c', Route11 ='c')) %>%
  janitor::clean_names() %>%
select(line, station_name, station_latitude, station_longitude, entry, exit_only, vending,
         entrance_type, ada, starts_with("route")) %>%
  mutate(entry= ifelse(entry== 'YES', TRUE, FALSE))
```

####Explanation
Currently the data set contains the variables: line, station name, station latitude, station longitude, routes served (1-11), entry, vending, entrance type, as well as if the station is ADA compliant. In order to get to this data set, it was essential to clean the variable names from the way they were converted in order to be interpretable for R. After this, the variables that were intended to be analysed were 'selected' from the data set to be used. Lastly, the variable entry was altered in order have it read as a logic variable rather than a character
variable, thus yes was interpreted as True, and no as False. The data is not currently 'tidy' at this point as route is still appearing as NA due to its format and thus unusable until it is altered. There are currently 20 columns and 1,868 rows in the current dataset.

####Distinct Stations
```{r}
nyc_transit%>%
  select(station_name,line) %>%
distinct
```
There are 465 distinct stations.

####ADA Compliance
```{r}
nyc_transit %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```
Of the 465 distinct stations, 84 are ADA compliant.

####Proportion of Allowed Entrances
```{r}
nyc_transit %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```
 The proportion of entrances/exits without vending allowed entrance is 0.377.

####Serve 'A' Train
```{r}
nyc_transit %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct
```
60 distinct stations serve the A train.

####Serve 'A' Train and ADA compliant
```{r}
nyc_transit %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```
Of the 60 distinct stations that serve the A train, 17 are ADA compliant.

##Problem 2

###Mr. Trash Wheel Data Set Cleaning
```{r}
mr_trash_wheel = 
  read_excel("Trash Wheel Collection Data.xlsx", 
    sheet = "Mr. Trash Wheel", col_names= TRUE) %>%
janitor::clean_names() %>%
  select(-x15, -x16) %>%
  drop_na(dumpster) %>%
  mutate(
    year = as.numeric(year),
  sports_balls= round(sports_balls),
  trash_wheel_data= "mrt" ) %>%
  select(trash_wheel_data, everything()) 
view(mr_trash_wheel)
skimr::skim(mr_trash_wheel)
```

###Professor Trash Wheel Data Set Cleaning
```{r}
prof_trash_wheel = 
  read_excel("Trash Wheel Collection Data.xlsx", 
    sheet = "Professor Trash Wheel", range= "A2:M96") %>%
janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    as.numeric(year),
  trash_wheel_data= "proft") %>%
  select(trash_wheel_data, everything()) 
view(prof_trash_wheel)
skimr::skim(prof_trash_wheel)
```

###Combining Mr. and Professor Trash Wheel Data Sets
```{r}
trash_wd_tidy = bind_rows(mr_trash_wheel, prof_trash_wheel) %>% 
  janitor:: clean_names() %>% 
  select(-as_numeric_year)
view(trash_wd_tidy)
skimr:: skim(trash_wd_tidy)
```

####Explanation
In this created data set, 'trash_wd_tidy', formed from 'mr_trash_wheel' and 'prof_trash_wheel', there are 641 observations and 15 variables. 547 of the variables were
from the 'mr_trash_wheel' data set while 94 were from the 'prof_trash_wheel' data. Some key variables include: 'weight_tons', 'volume_cubic_yards', as well as 'dumpster'. An additional variable called 'trash_wheel_data was used in order to differentiate the two data sets when combining together. However, in this updated data set, only 'mr_trash_wheel' contained the variable for sports_balls. This was not a variable in the updated 'prof_trash_wheels', thus it returned NA results in the 'trash_wd_tidy' data set created.

####Total Weight of Trash Collected: Professor Trash Wheel
```{r}
sum(prof_trash_wheel$weight_tons)
```
The total weight of trash collected by Professor Trash Wheel was 190.12 tons.

####Number of Sports Balls: Mr. Trash Wheel
```{r}
sum(mr_trash_wheel$sports_balls)
sum(subset(mr_trash_wheel, year== '2020' )$sports_balls)
```
The total number of balls collected by Mr. Trash Wheel in 2020 was 856 balls.

##Problem 3

###Pols-Month Data Cleaning
```{r}
fte_pols = read_csv("/Users/mayurialbal/Documents/Data Science/Homework/HW2/HW2/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, c("year", "month", "day"), sep = "-") %>% 
  mutate( 
  month = recode(month,
      '01' = "january", '02'= "february", '03'= "march", '04'= "april", '05'= "may",
      '06'= "june", '07'= "july", '08'= "august", '09'= "september", '10'= "october", 
      '11' = "november", '12' = "december"),
  prez_dem= recode(prez_dem, "1"= "democrat" ),
  prez_gop= recode (prez_gop, "1"= "republican")) %>% 
  pivot_longer(cols= c("prez_dem", "prez_gop"),
               values_to= "president") %>% 
  drop_na ("president") %>% 
  select(-day,-name)
view(fte_pols)
```

###SNP Data Cleaning
```{r}
fte_snp = read_csv("/Users/mayurialbal/Documents/Data Science/Homework/HW2/HW2/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
     date= as.Date(date,format= c("%m/%d/%y")),
     date= as.Date(ifelse(date > "2015-07-01", format(date,"19%y-%m-%d"), format(date)))) %>%
  separate(date, c("year", "month", "day"), sep = "-") %>% 
mutate(
  month = recode(month,
      '01' = "january", '02'= "february", '03'= "march", '04'= "april", '05'= "may",
      '06'= "june", '07'= "july", '08'= "august", '09'= "september", '10'= "october", 
      '11' = "november", '12' = "december")) %>%
  arrange(year, month) %>% 
  select(-day) %>% 
view(fte_snp)
```

###Cleaning the Unemployment Data
```{r}
fte_unem = read_csv("/Users/mayurialbal/Documents/Data Science/Homework/HW2/HW2/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
pivot_longer(jan:dec, names_to = "month")
  
fte_unem$year <- as.character(fte_unem$year)
view(fte_unem)
```

###Merge fte_pols and fte_snp, as well as fte_unem
```{r}
fte_pols_snp=
  left_join(fte_pols, fte_snp, by= c("year", "month"))
view(fte_pols_snp)
```

```{r}
fte_merge=
  left_join(fte_pols_snp, fte_unem, by= c("year", "month"))
view(fte_merge)
```

####Explanation
FTE_POLS:
Within the "fte_pols" data set there were 817 observations and 9 variables, after completion
of cleaning. The years in the data set ranged from 1947-2015. The variables found in
the data set included: 'year','month','gov_gop','gov_dem','sen_gop', 'sen_dem', 'rep_gop', 'rep_dem' as well as 'president'. This data set contained data regarding the amount
of national politicians who were either democratic or republican at any point in the data
range.

FTE_SNP:
For the data set "fte_snp" there were 787 observations and 3 variables, after cleaning of the data. The years in the data set ranged from 1950-2015. The variables in the data set were: 'year', 'month', and 'close'. This data set contained information regarding stock market
index closing values on the attributed date in this range.

FTE_UNEM
The data set "fte_unem" had 816 observations and 3 variables, after cleaning of data. The years in the data set ranged from 1948-2015. The variables in the data set were: 'year, 'month', and 'value'. The 'value' variable was the reported percentage of unemployment in that given month within the specified year.

FTE_MERGE
This data set was a combination of "fte_pols" and "fte_snp", with the later merging of 'fte_unem. It contained 817 observations and 11 variables. The variables contained in the set were: 'year','month','gov_gop','gov_dem','sen_gop', 'sen_dem', 'rep_gop', 'rep_dem', 'president', 'close', and 'value. The data set ranged from 1947-2015, however the data for 
variables 'close' ranged from 1950-2015, and 1948-2015 for 'value'. This data set provided a combined set of all the selected sets to allow for an understanding of governmental political parties in power and their relation to the stock market as well as unemployment rates.
